---
title: "Binding Values to Symbol"
author: "eemilSinclair"
date: "5/18/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Binding

When R tries to bind a value to a symbol it searches through a series of environments to find the appropriate value. The order is:

1. Global enviroment
2. Namespaces of each package on the list

```{r namespace}
search()
```

* Global Enviroment is the first package in the list
* Base package is always the last
* Order matters!
* Users **can configure** which packages get loaded on startup
* Loading packages with 'library' the namespace of the package gets put into second position of the list
* There are **separte** namespaces for functions and non-functions. So you can have a vector and a function named 'c'

## Environments
Is a collection of symbol/value pairs. They can have parent environments, and a parent can have multiple children. The only enviroment without a parent is the empty environment.

A function plus an enviroment equals a **closure** or **function closure**.

## Scoping Rules

* Function arguments: available at the definition of the function.
* The are other symbols not defined by function arguments.

So we have alternatives:

### 1. Lexical or Static Scoping:

Useful for statistic calculations. The value of free variables is searched in the enviroment the function was defined in.
```{R lexical_scoping}
  f <- function(x, y){
    
    #'z' is a free variable. Free variables are not local variales nor formal arguments.
    x^2 + y / z
  }
```

1. A free value is first searched for in the environment where the function was defined (can be the global).
2. If it's not there, it will look in the parent environment, until it hits the top-level environment.
3. Then it goes to the next enviroment on the list and its parents. 
4. Not found? Looks in the empty environment.
5. Not found yet? **ERROR**

**KEY DIFFERENCE WITH OTHER LANGUAGES**
If a function is defined inside another function, the environment of the first function is the second one.
```{R nested_functions_env}

make.power <- function(n){
  pow <- function(x){
    x^n
  }
  pow
}

cube <- make.power(3)
square <- make.power(2)
cube(4)
square(5)
```

Exploring a function closure:
```{R closure}
ls(environment(cube))
get("n",environment(cube))

ls(environment(square))
get("n",environment(square))
```

### 2. Dynamic Scoping
```{R dynamic_scoping}
y <- 10

f <- function(x){
  #Local variable
  y <- 2
  
  #y is in the same calling environment (2).
  y^2 + g(x)
}

g <- function(x){
  
  #y is called from the function's parent frame (10).
  x*y
}

f(3)

```

When a function is defined in the global environment, and it is called from the global environment, then the defining and the calling environment are the same. This is not dynamic scoping.
