---
title: "Control Structures 2"
author: "eemilSinclair"
date: "5/18/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## While Loops

Example of while loops. There are useful because they are easy to read, but if the counter variable is not handled correctly, they can become infinite quite fast.

```{r while}
count <- 0
while(count < 10){
  print(count)
  count <- count + 1
}
```

Multiple conditions on a while loop. Conditions are evaluated from left to right:
```{r multiple_cond_while}

z <- 5
while(z >= 3 && z < 10){
  print(z)
  
  coin <- rbinom(1,1,0.5)
  print(c("coin",coin))
  
  if(coin == 1){
    z <- z +1
  }else{
    z <- z - 1
  }
}

```

## Repeat, next, break

Repeat initiates an infinite loop. We can exit using 'break':
```{r repeat}

x0 <- 1
tol <- 1e-8

repeat{
  x1 <- 0.123443
  print(x1)
  if(abs(x1 - x0) < tol){
    break
  }else{
    x0 <- x1
  }
}

```

Next allows skiping to the next iteration. 'return' exits the fuction and return a given value.
```{r next}

for(i in 1:100){
  if(i <= 20){
    ##Skip the first 20 iterations...
    next
  }
  print(i)
}
```
