---
title: "Functions"
author: "eemilSinclair"
date: "5/18/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Functions

Functions are first class objects. You can pass functions as arguments to other functions, as well as have nested functions. The return expression is the last evaluated line inside the function.

Functions have named arguments:

* Formal argument: included in the function definition.
* The 'formals' function returns a list of the formal aguments in a function.
* Not every function in R makes use of the formal arguments.
* Function arguments might be missing or have default values.

R function arguments can be matched by **position** or by **name**. When an argument is matched by name, it is taken out of the argument list and the next unnamed arguments are matched in the order they're listed.

### Partial Matching

The order of operations when given an argument is:

1. Exact match
2. Partial match
3. Position match

```{r formals}
formals(lm);

```

```{r args}
args(lm);

```

Hello World!
```{r hello_world}
add2 <- function(x, y){
  x + y
}
```

Return the subset of a vector above values:
```{r above_10}

above10 <- function(x){
  use <- x > 10
  x[use]
}

above <- function(x,n){
  use <- x > n
  x[use]
}


abovedefault <- function(x,n=25){
  use <- x > n
  x[use]
}
```

Function to extract the mean of each column. Adding the removeNA parameter of the function to remove NA values on the original dataframe.
```{r column_mean, echo=FALSE, message=FALSE}

columnmean <- function(y, removeNA = TRUE){
  nc <- ncol(y)
  means <- numeric(nc)
  
  for(i in 1:nc){
    means[i] <- mean(y[,i], na.rm = removeNA)
  }
  
  means
}

```

## Lazy Evaluation

Evalute functions only when they are needed. For example, 'b' is never evaluated, so there will not be an error because 2 gets pisitionally matched at 'a'.
```{R lazyev}
f <- function(a,b){
  a^2
}

f(2)
```

```{R lazyev2}
f <- function(a,b){
  print(a)
  
  #Commenting as this will not work!
  #print(b)
}

f(2)
```

# The '...' Argument

Indicates a vaariable number of arguments. Used to extend other functions without writing the whole argument list of the original.
```{r dotdotdot}

#Extending the plot function
myplot <- function(x,y,type=1,...){
  plot(x,y,type=type,...)
}

#Use in generic functions to pass extra arguments to methods
#No example so far...

# When number of arguments can't be known in advance:
args(paste)

args(cat)

# All the arguments coming after '...' must be explicitly matched
args(paste)
paste("a","b",se=":")
```