---
title: "Loop Functions"
author: "eemilSinclair"
date: "5/22/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loop Functions

### lapply Function.

Loop over a list of objects and apply that function on each element of the list. 3 arguments: the list, a function to apply, and other functions by ...

If X is not a list, X is coerced to a list using as.list

```{r lapply1}
x <- list(a=1:5, b=rnorm(10))
y <-lapply(x, mean)
print(y)
class(y)
```
```{r lapply2}
x <- 1:4
y <-lapply(x, runif, min=0, max=10)
print(y)
class(y)
```

```{r lapply3}
x <- list(a = matrix(1:4,2,2), b= matrix(1:6,3,2))
print(x)
#Anonymous function elt
lapply(x, function(elt)elt[,1])
```

### Sapply Function.

Loop, simplifies the result of lapply. 
* If the result is a list where each element is of length 1, a vector is returned.
* If the result is a list of vectors of the same length > 1, a matrix is returned.
* Otherwise, a list is returned.
```{r sapply}
x <- list(a=1:5, b=rnorm(10))
y <-sapply(x, mean)
print(y)
class(y)
```

### Apply Function.

Loop, evaluate a function over the margins of an array.
* Apply a function to the rows or columns of a matrix
* It can be used on general arrays.
```{r apply}
x <- matrix(rnorm(200),20,10)

#Mean of COLUMNS
y <-apply(x, 2, mean)
print(attributes(x))
print(y)
class(y)

#Sum of each ROW of the matrix
z <-apply(x, 1, sum)
print(z)
class(z)
```
### Mapply Function.

Loop, multivariant function.

```{r mapply}
x <- mapply(rep, 1:4, 4:1)
print(x)
class(x)
```

### Tapply Function.

Over subsets of vector. Application example: men and women. We need to have a factor variable to differentiate from them. 
```{r tapply}
x <- c(rnorm(10),runif(10),rnorm(10,1))
print(x)
f <- gl(3,10)
print(f)
y <- tapply(x, f, mean,simplify=TRUE)
print(y)
class(y)
z <- tapply(x, f, mean,simplify=FALSE)
print(z)
class(z)
```

### Split Function.

Takes a vector or other objects and splits them based on a factor or list of factors. Always returns a list.
```{r split}
#Divide one single vector X into three parts, defined by F...
x <- c(rnorm(10),runif(10),rnorm(10,1))
print(x)
f <- gl(3,10)
print(f)
y <- split(x,f)
print(y)
class(y)

lapply(split(x,f), mean)
```

```{r split2}
library(datasets)

#Split air quality dataset by month...
s<-split(airquality,airquality$Month)
print(s)
class(s)

#Subset by Ozone, Solar Radiation, Wind. Get the means of each variable.
x <- lapply(s,function(x)colMeans(x[,c("Ozone", "Solar.R","Wind")]))
print(x)
class(x)

#Get the means of each variable.
y <-sapply(s,function(x)colMeans(x[,c("Ozone", "Solar.R","Wind")], na.rm=TRUE))
str(y)
print(y)
class(y)

#Check the interaction() function...
```